Full Dissertation

### 2024-04-06 Sections/Introduction.qmd
render full in pdf and markdown

---
title: Full Dissertation
subtitle: An instance-based model account of the benefits of varied practice in visuomotor skill
date: 04/04/24
#page-layout: full
code-fold: true
code-tools: true
toc: true
toc-location: body
toc-title: Contents
execute: 
  warning: false
  eval: true
format:
  html: default
  hugo-md:
    echo: false
    output: false
    html-math-method: mathjax
    output-file: diss_full_hugo.md
  gfm:
    echo: true
    output-file: diss_full_gfm.md
  hikmah-manuscript-pdf:
    echo: false
    output-file: "tg_full_disser.pdf"
    mainfont: "Linux Libertine O"
    mainfontoptions:
      - "Numbers=Proportional"
      - "Numbers=OldStyle"
    mathfont: "Libertinus Math"
editor: 
  markdown: 
    wrap: 72
prefer-html: true
---

@@

# Project 1

____________________________________________________
[pdf of the journal
article](Assets/Gorman_Goldstone_2022_Instance-based_model_varied_practice.pdf){target="_blank"}\
[Link to online version of journal
article](https://www.sciencedirect.com/science/article/abs/pii/S0010028522000299){target="_blank"}

```{r}

pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,
  brms,tidybayes, rstanarm,emmeans,broom,bayestestR,
  stringr, here,conflicted, patchwork, knitr, cowplot, RColorBrewer,data.table,gt,
  ggh4x,htmltools,future,furrr,ez)
#options(brms.backend="cmdstanr",mc.cores=4)
#options(digits=2, scipen=999, dplyr.summarise.inform=FALSE, kableExtra.auto_format=FALSE)
walk(c("brms","dplyr","bayestestR","here"), conflict_prefer_all, quiet = TRUE)
walk(c("Display_Functions","fun_alm","fun_indv_fit","fun_model", "prep_model_data","org_functions"), ~source(here::here(paste0("Functions/", .x, ".R"))))
#conflicts_prefer(tidybayes::rstudent_t)

# pacman::p_load(dplyr,purrr,tidyr,ggplot2, data.table, here, patchwork, conflicted, 
#                stringr,future,furrr, knitr, reactable,ggstance, htmltools,
#                ggdist,ggh4x,brms,tidybayes,emmeans,bayestestR, gt)

#pacman::p_load(papaja, tinytex, RColorBrewer, kableExtra, cowplot)
source(here::here('Functions/IGAS_ProcessFunctions.R'))

#options(brms.backend="cmdstanr",mc.cores=4)
initial_contrasts <- options("contrasts")$contrasts
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE, contrasts = c("contr.sum", "contr.poly"))
defaultContrasts = options()$contrasts

library(tidybayes)

theme_set(theme_classic())
# load the processed data from experiment 1 and 2
e1 <- readRDS(here::here("data/igas_e1_cleanedData-final.rds")) %>% mutate(initialVelocityX=X_Velocity,initialVelocityY=Y_Velocity,stageInt=as.numeric(as.character(experimentStage)))
e2<- readRDS(here::here('data/igas_e2_cleanedData-final.rds')) %>% mutate(initialVelocityX=X_Velocity,initialVelocityY=Y_Velocity)
# load subject similarity data - computed with the IGAS model in 'IGAS-SimModel.R'
e2_sim <- readRDS(here::here('data/IGAS_Similarity-Performance.rds'))

# options(contrasts = c("contr.sum", "contr.poly"), kableExtra.auto_format=FALSE)
# defaultContrasts = options()$contrasts
# theme_set(theme_classic())

dodge <- position_dodge(width = 0.9)
e2GrpPos <- c("400","500","625","675","800","900")
e2Grp <- paste("Constant","Constant", "Constant","Constant","Constant","Constant", "Varied")
e2Labels <- paste(c("400\n Constant","500\n Constant","625\n Constant","675\n Constant",
                   "800\n Constant","900\n Constant","500-800\n Varied"),sep="")

e1Pos <- c("610","760","835","910")
e1Var <- paste("Varied Train Position","Constant Train Position", "Novel Position", "Varied Training Position")
e1Labels<- paste(c("610\n Varied Trained","760\n Constant Trained","835\n Novel Location","910\n Varied Trained"),sep="")
```

# Abstract

Exposing learners to variability during training has been demonstrated to improve performance in subsequent transfer testing. Such variability benefits are often accounted for by assuming that learners are developing some general task schema or structure. However much of this research has neglected to account for differences in similarity between varied and constant training conditions. In a between-groups manipulation, we trained participants on a simple projectile launching task, with either varied or constant conditions. We replicate previous findings showing a transfer advantage of varied over constant training. Furthermore, we show that a standard similarity model is insufficient to account for the benefits of variation, but, if the model is adjusted to assume that varied learners are tuned towards a broader generalization gradient, then a similarity-based model is sufficient to explain the observed benefits of variation. Our results therefore suggest that some variability benefits can be accommodated within instance-based models without positing the learning of some schemata or structure.

# Introduction

The past century of research on human learning has produced ample evidence that although learners can improve at almost any task, such improvements are often specific to the trained task, with unreliable or even nonexistent transfer to novel tasks or conditions [@barnettWhenWhereWe2002; @dettermanCaseProsecutionTransfer1993]. Such transfer challenges are of noteworthy practical relevance, given that educators, trainers, and rehabilitators typically intend for their students to be able to apply what they have learned to new situations. It is therefore important to better understand the factors that influence transfer, and to develop cognitive models that can predict when transfer is likely to occur. The factor of interest to the present investigation is variation during training. Our experiments add to the longstanding empirical investigation of the controversial relationship between training variation, and subsequent transfer. We also offer a novel explanation for such results in the form of an instance-based model that accounts for the benefits of variation in simple terms of psychological similarity. We first review the relevant concepts and literature.

## Similarity and instance-based approaches to transfer of learning

Notions of similarity have long played a central role in many prominent models of generalization of learning, as well as in the longstanding theoretical issue of whether learners abstract an aggregate, summary representation, or if they simply store individual instances. Early models of learning often assumed that discrete experiences with some task or category were not stored individually in memory, but instead promoted the formation of a summary representation, often referred to as a prototype or schema, and that exposure to novel examples would then prompt the retrieval of whichever preexisting prototype was most similar [@posnerGenesisAbstractIdeas1968]. Prototype models were later challenged by the success of instance-based or exemplar models -- which were shown to provide an account of generalization as good or better than prototype models, with the advantage of not assuming the explicit construction of an internal prototype [@estesClassificationCognition1994; @hintzmanMINERVASimulationModel1984; @medinContextTheoryClassification1978;
@nosofskyAttentionSimilarityIdentificationcategorization1986 ]. Instance-based models assume that learners encode each experience with a task as a separate instance/exemplar/trace, and that each encoded trace is in turn compared against novel stimuli. As the number of stored instances increases, so does the likelihood that some previously stored instance will be retrieved to aid in the performance of a novel task. Stored instances are retrieved in the context of novel stimuli or tasks if they are sufficiently similar, thus suggesting that the process of computing similarity is of central importance to generalization.

Similarity, defined in this literature as a function of psychological distance between instances or categories, has provided a successful account of generalization across numerous tasks and domains. In an influential study demonstrating an ordinal similarity effect, experimenters employed a numerosity judgment task in which participants quickly report the number of dots flashed on a screen. Performance (in terms of response times to new patterns) on novel dot configurations varied as an inverse function of their similarity to previously trained dot configurations @palmeriExemplarSimilarityDevelopment1997. That is, performance was better on novel configurations moderately similar to trained configurations than to configurations with low-similarity, and also better on low-similarity configurations than to even less similar, unrelated configurations. Instance-based approaches have had some success accounting for performance in certain sub-domains of motor learning [@cohenWhereGraspsAre2004; @crumpEpisodicContributionsSequential2010; @meighWhatMemoryRepresentation2018; @poldrackRelationshipSkillLearning1999; @wifallReachingResponseSelection2017]. @crumpEpisodicContributionsSequential2010 trained participants to type words on an unfamiliar keyboard, while constraining the letters composing the training words to a pre-specified letter set. Following training, typing speed was tested on previously experienced words composed of previously experienced letters; novel words composed of letters from the trained letter set; and novel words composed of letters from an untrained letter set. Consistent with an instance-based account, transfer performance was graded such that participants were fastest at typing the words they had previously trained on, followed by novel words composed of letters they had trained on, and slowest performance for new words composed of untrained letters.

## The effect of training variability on transfer

While similarity-based models account for transfer by the degree of similarity between previous and new experiences, a largely separate body of research has focused on improving transfer by manipulating characteristics of the initial training stage. Such characteristics have included training difficulty, spacing, temporal order, feedback schedules, and the primary focus of the current work -- variability of training examples.

Research on the effects of varied training typically compares participants trained under constant, or minimal variability conditions to those trained from a variety of examples or conditions [@czyzVariabilityPracticeInformation2021; @soderstromLearningPerformanceIntegrative2015]. Varied training has been shown to influence learning in myriad domains including categorization of simple stimuli [@hahnEffectsCategoryDiversity2005; @maddoxStimulusRangeDiscontinuity2011; @posnerGenesisAbstractIdeas1968], complex categorization [@nosofskyModelguidedSearchOptimal2018], language learning [@jonesDensityDistinctivenessEarly2020; @perryLearnLocallyThink2010; @twomeyAllRightNoises2018; @wonnacottInputEffectsAcquisition2012], anagram completion [@goodeSuperiorityVariableRepeated2008], trajectory extrapolation [@fulvioTaskSpecificResponseStrategy2014], task switching [@sabahWhenLessMore2019], associative learning [@leeEvidentialDiversityIncreases2019], visual search [@georgeStimulusVariabilityTask2021; @gonzalezDiversityTrainingEnhances2011; @kelleyLearningAttendEffects2009], voice identity learning [@lavanEffectsHighVariability2019], simple motor learning [@braunMotorTaskVariation2009; @kerrSpecificVariedPractice1978; @rollerVariablePracticeLenses2001; @willeyLimitedGeneralizationVaried2018], sports training  [@greenPracticeVariabilityTransfer1995a; @northEffectConsistentVaried2019], and training on a complex video game [@seowTransferEffectsVaried2019].

Training variation has received a particularly large amount of attention within the domain of visuomotor skill learning. Much of this research has been influenced by the work of @schmidtSchemaTheoryDiscrete1975, who proposed a schema-based account of motor learning as an attempt to address the longstanding problem of how novel movements are produced. According to Schema Theory, learners possess general motor programs for classes of movements (e.g. throwing a ball with an underhand movement), as well as schema rules that determine how a motor program is parameterized or scaled for a particular movement. Schema theory predicts that varied training results in the formation of a more general schema-rule, which can allow for transfer to novel movements within a given movement class. Experiments that test this hypothesis are often designed to compare the transfer performance of a constant-trained group against that of a varied-trained group. Both groups train on the same task, but the varied group practices from multiple levels of a task-relevant dimension that remains invariant for the constant group. For example, investigators might train two groups of participants to throw a projectile at a target, with a constant group that throws from a single location, and a varied group that throws from multiple locations. Both groups are then tested from novel locations. Empirically observed benefits of the varied-trained group are then attributed to the variation they received during training, a finding observed in numerous studies [@catalanoDistantTransferCoincident1984a; @chuaPracticeVariabilityPromotes2019; @goodwinEffectDifferentQuantities1998; @kerrSpecificVariedPractice1978; @wulfEffectTypePractice1991], and the benefits of this variation are typically thought to be mediated by the development of a more general schema for the throwing motion.

Of course, the relationship between training variability and transfer is unlikely to be a simple function wherein increased variation is always beneficial. Numerous studies have found null, or in some cases negative effects of training variation [@deloshExtrapolationSineQua1997; @sinkeviciuteRoleInputVariability2019; @wrisbergVariabilityPracticeHypothesis1987], and many more have suggested that the benefits of variability may depend on additional factors such as prior task experience, the order of training trials, or the type of transfer being measured [@bernikerEffectsTrainingBreadth2014; @braithwaiteEffectsVariationPrior2015; @hahnEffectsCategoryDiversity2005; @lavanEffectsHighVariability2019; @northEffectConsistentVaried2019; @sadakataIndividualAptitudeMandarin2014; @zamanPerceptualVariabilityImplications2021].

## Issues with Previous Research

Although the benefits of training variation in visuomotor skill learning have been observed many times, null findings have also been repeatedly found, leading some researchers to question the veracity of the variability of practice hypothesis [@newellSchemaTheory19752003; @vanrossumSchmidtSchemaTheory1990]. Critics have also pointed out that investigations of the effects of training variability, of the sort described above, often fail to control for the effect of similarity between training and testing conditions. For training tasks in which participants have numerous degrees of freedom (e.g. projectile throwing tasks where participants control the x and y velocity of the projectile), varied groups are likely to experience a wider range of the task space over the course of their training (e.g. more unique combinations of x and y velocities).
Experimenters may attempt to account for this possibility by ensuring that the training location(s) of the varied and constant groups are an equal distance away from the eventual transfer locations, such that their training throws are, on average, equally similar to throws that would lead to good performance at the transfer locations. However, even this level of experimental control may still be insufficient to rule out
the effect of similarity on transfer. Given that psychological similarity is typically best described as either a Gaussian or exponentially decaying function of psychological distance [@ennisMultidimensionalStochasticTheory1988; @ghahramaniGeneralizationLocalRemappings1996; @loganInstanceTheoryAutomatization1988; @nosofskySimilarityScalingCognitive1992; @shepardUniversalLawGeneralization1987; @thoroughmanRapidReshapingHuman2005 ], it is plausible that a subset of the most similar training instances could have a disproportionate impact on generalization to transfer conditions, even if the average distance between training and transfer conditions is identical between groups. @fig-toy-model1 demonstrates the consequences of a generalization gradient that drops off as a Gaussian function of distance from training, as compared to a linear drop-off.

```{r}
#| label: fig-toy-model1
#| fig-cap: Left panel- Generalization predicted from a simple model that assumes a linear generalization function. A varied group (red vertical lines indicate the 2 training locations) trained from positions 400 and 800, and a constant group (blue vertical line), trained from position 600. Right panel- if a Gaussian generalization function is assumed, then varied training (400, 800) is predicted to result in better generalization to positions close to 400 and 800 than does constant training at 600. (For interpretation of the references to color in this figure legend, the reader is referred to the web version of this article.)

p=2
c<- .0002
simdat <- data.frame(x=rep(seq(200,1000),3),condit=c(rep("varied",1602),rep("constant",801)),
                     train.position=c(rep(400,801),rep(800,801),rep(600,801)),c=.0002,p=2) %>%
                     mutate(plotjitter=ifelse(condit=="varied",0,7),
                            linScale=ifelse(condit=="varied",980,1000),
                            genGauss=exp(-c*(abs((x-train.position)^p))),
                            genLinear=1000-abs(x-train.position)+plotjitter) %>% 
  #group_by(condit) %>% mutate(scaleLinear=(genLinear-min(genLinear))/(max(genLinear)-min(genLinear))) 
  group_by(x,condit) %>%
  reframe(genGauss=mean(genGauss),genLinear=mean(genLinear)/linScale,.groups = 'keep')
colorVec=c("darkblue","darkred")
plotSpecs <- list(geom_line(alpha=.7,size=.4),scale_color_manual(values=colorVec),
                  geom_vline(alpha=.55,xintercept = c(400,800),color=colorVec[2]),
                  geom_vline(alpha=.55,xintercept = c(600),color=colorVec[1]),
                  ylim(c(0,1.05)),
                  #xlim(c(250,950)),
                  scale_x_continuous(breaks=seq(200,1000,by=200)),
                  xlab("Test Stimulus"),
                  annotate(geom="text",x=447,y=1.05,label="Varied",size=3.1,fontface="plain"),
                  annotate(geom="text",x=450,y=1.02,label="Training",size=3.1,fontface="plain"),
                  annotate(geom="text",x=659,y=1.05,label="Constant",size=3.1,fontface="plain"),
                  annotate(geom="text",x=657,y=1.02,label="Training",size=3.1,fontface="plain"),
                  annotate(geom="text",x=847,y=1.05,label="Varied",size=3.1,fontface="plain"),
                  annotate(geom="text",x=850,y=1.02,label="Training",size=3.1,fontface="plain"),
                  theme(panel.border = element_rect(colour = "black", fill=NA, linewidth=1),
                        legend.position="none"))

ip1 <- simdat  %>% ggplot(aes(x,y=genGauss,group=condit,col=condit))+plotSpecs+ylab("")
ip2 <- simdat %>%  ggplot(aes(x,y=genLinear,group=condit,col=condit))+plotSpecs+ylab("Amount of Generalization")

plot_grid(ip1,ip2,ncol=2,rel_heights=c(1))

```

In addition to largely overlooking the potential for non-linear generalization to confound interpretations of training manipulations, the visuomotor skill learning literature also rarely considers alternatives to schema representations [@chamberlinMemoryRepresentationMotor1992]. Although schema-theory remains influential within certain literatures, instance or exemplar-based models have accounted for human behavior across myriad domains [@jamiesonInstanceTheoryDomaingeneral2022; @loganInstanceTheoryAttention2002a]. As mentioned above, instance based accounts have been shown to perform well on a variety of different tasks with motoric components [@crumpEpisodicContributionsSequential2010;@gandolfoMotorLearningField1996a;  @meighWhatMemoryRepresentation2018; @rosenbaumPlanningReachesEvaluating1995; @vandamMappingShapeVisuomotor2015]. However, such accounts have received little attention within the subdomain of visuomotor skill learning focused on the benefits of varied training.

The present work examines whether the commonly observed benefits of varied training can be accounted for by between-group differences in similarity between training and testing throws. We first attempt to replicate previous work finding an advantage of varied training over constant training in a projectile launching task. We then examine the extent to which this advantage can be explained by an instance-based similarity model.

# Experiment 1

## Methods

### Sample Size Estimation

To obtain an independent estimate of effect size, we identified previous
investigations which included between-subjects contrasts of varied and
constant conditions following training on an accuracy based projectile
launching task [@chuaPracticeVariabilityPromotes2019; @goodwinEffectDifferentQuantities1998; @kerrSpecificVariedPractice1978; @wulfEffectTypePractice1991]. We then averaged effects across these studies,
yielding a Cohens f =.43. The GPower 3.1 software package [@faulStatisticalPowerAnalyses2009],
2009) was then used to determine that a power of 80% requires a sample
size of at least 23 participants per condition. All experiments reported
in the present manuscript exceed this minimum number of participants per
condition.

### Participants

Participants were recruited from an undergraduate population that is 63%
female and consists almost entirely of individuals aged 18-22 years. A
total of 110 Indiana University psychology students participated in
Experiment 1. We subsequently excluded 34 participants poor performance
at one of the dependent measures of the task (2.5-3 standard deviations
worse than the median subject at the task) or for displaying a pattern
of responses that was clearly indicative of a lack of engagement with
the task (e.g. simply dropping the ball on each trial rather than
throwing it at the target), or for reporting that they completed the
experiment on a phone or tablet device, despite the instructions not to
use one of these devices. A total of 74 participants were retained for
the final analyses, 35 in the varied group and 39 in the constant group.

### Task

The experimental task was programmed in JavaScript, using packages from
the Phaser physics engine (https://phaser.io) and the jsPsych library
(de Leeuw, 2015). The stimuli, presented on a black background,
consisted of a circular blue ball -- controlled by the participant via
the mouse or trackpad cursor; a rectangular green target; a red
rectangular barrier located between the ball and the target; and an
orange square within which the participant could control the ball before
releasing it in a throw towards the target. Because the task was
administered online, the absolute distance between stimuli could vary
depending on the size of the computer monitor being used, but the
relative distance between the stimuli was held constant. Likewise, the
distance between the center of the target, and the training and testing
locations was scaled such that relative distances were preserved
regardless of screen size. For the sake of brevity, subsequent mentions
of this relative distance between stimuli, or the position where the
ball landed in relation to the center of the target, will be referred to
simply as distance. @fig-IGAS_Methods displays the layout of the task, as
it would appear to a participant at the start of a trial, with the ball
appearing in the center of the orange square. Using a mouse or trackpad,
participants click down on the ball to take control of the ball,
connecting the movement of the ball to the movement of the cursor.
Participants can then "wind up" the ball by dragging it (within the
confines of the orange square) and then launch the ball by releasing the
cursor. If the ball does not land on the target, participants are
presented with feedback in red text at the top right of the screen, on
how many units away they were from the center of the target. If the ball
was thrown outside of the boundary of the screen participants are given
feedback as to how far away from the target center the ball would have
been if it had continued its trajectory. If the ball strikes the barrier
(from the side or by landing on top), feedback is presented telling
participants to avoid hitting the barrier. If participants drag the ball
outside of the orange square before releasing it, the trial terminates,
and they are reminded to release the ball within the orange square. If
the ball lands on the target, feedback is presented in green text,
confirming that the target was hit, and presenting additional feedback
on how many units away the ball was from the exact center of the target.

[Link to abbrevaited example of
task](https://pcl.sitehost.iu.edu/tg/demos/igas_expt1_demo.html){target="_blank"}.

```{r}
#| label: fig-IGAS_Methods
#| fig.cap: The stimuli of the task consisted of a blue ball, which the participants would launch at the green target, while avoiding the red barrier. On each trial, the ball would appear in the center of the orange square, with the position of the orange square varying between experimental conditions. Participants were constrained to release the ball within the square

mf <- cowplot::ggdraw()+cowplot::draw_image(here::here("Assets/methodsFig1.png"),hjust=0)+theme(plot.margin = margin(0, 0, 0, 0))
plot_grid(mf,ncol=1)

```

## Results

## Data Processing and Statistical Packages

To prepare the data, we first removed trials that were not easily
interpretable as performance indicators in our task. Removed trials
included: 1) those in which participants dragged the ball outside of the
orange starting box without releasing it, 2) trials in which
participants clicked on the ball, and then immediately released it,
causing the ball to drop straight down, 3) outlier trials in which the
ball was thrown more than 2.5 standard deviations further than the
average throw (calculated separately for each throwing position), and 4)
trials in which the ball struck the barrier. The primary measure of
performance used in all analyses was the absolute distance away from the
center of the target. The absolute distance was calculated on every
trial, and then averaged within each subject to yield a single
performance score, for each position. A consistent pattern across
training and testing phases in both experiments was for participants to
perform worse from throwing positions further away from the target -- a
pattern which we refer to as the difficulty of the positions. However,
there were no interactions between throwing position and training
conditions, allowing us to collapse across positions in cases where
contrasts for specific positions were not of interest. All data
processing and statistical analyses were performed in R version 4.03 (R
Core Team, 2020). ANOVAs for group comparisons were performed using the
rstatix package (Kassambara, 2021)****.

## Training Phase

@fig-IGAS_Training1 below shows aggregate training performance binned into three
stages representing the beginning, middle, and end of the training
phase. Because the two conditions trained from target distances that
were not equally difficult, it was not possible to directly compare
performance between conditions in the training phase. Our focus for the
training data analysis was instead to establish that participants did
improve their performance over the course of training, and to examine
whether there was any interaction between training stage and condition.
Descriptive statistics for the intermittent testing phase are provided
in the supplementary materials.

We performed an ANOVA comparison with stage as a within-group factor and
condition as between-group factor. The analysis revealed a significant
effect of training stage F(2,142)=62.4, p\<.001, $\eta^{2}_G$ = .17,
such that performance improved over the course of training There was no
significant effect of condition F(1,71)=1.42, p=.24, $\eta^{2}_G$ = .02,
and no significant interaction between condition and training stage,
F(2,142)=.10, p=.91, $\eta^{2}_G$ \< .01.

```{r}
#| label: fig-IGAS_Training1
#| fig.cap: Training performance for varied and constant participants binned into three stages. Shorter bars indicate better performance (ball landing closer to the center of the target). Error bars indicate standard error of the mean.

exp1TrainPosition <- e1 %>% filter(stage!="Transfer",mode==1) %>%ungroup() %>% 
  group_by(sbjCode,Group,conditType,trainHalf,positionX) %>% 
  summarise(MeanTargetDistance=mean(AbsDistFromCenter),.groups = 'keep')

exp1TrainPosition3 <- e1 %>% filter(stage!="Transfer",mode==1) %>%ungroup() %>% 
  group_by(sbjCode,Group,conditType,stage,positionX) %>% 
  summarise(MeanTargetDistance=mean(AbsDistFromCenter),.groups = 'keep')

exp1Train <- e1 %>% filter(stage!="Transfer",mode==1)  %>%
  group_by(sbjCode,Group,conditType,trainHalf) %>% 
  summarise(MeanTargetDistance=mean(AbsDistFromCenter),.groups = 'keep')

exp1Train3 <- e1 %>% filter(stage!="Transfer",mode==1)  %>%
  group_by(sbjCode,Group,conditType,stage) %>% 
  summarise(MeanTargetDistance=mean(AbsDistFromCenter),.groups = 'keep')

e1train2 <- exp1TrainPosition3 %>% ggplot(aes(x=positionX,y=MeanTargetDistance))+
  geom_bar(aes(group=stage,fill=stage),stat="summary",fun=mean,position=dodge)+
  facet_wrap(~conditType,ncol=2)+
  stat_summary(aes(x=positionX,group=stage),fun.data=mean_se,geom="errorbar",position=dodge,width=.8)+
  ylab("Mean Distance From Center Of Target")+
  xlab("Training Location(s)")+theme(plot.title = element_text(hjust = 0.5))+
  guides(fill=guide_legend(title="Training Stage"))+theme(legend.title.align=.25)

#plot_grid(title,e1train2,capt,ncol=1,rel_heights=c(.18,1,.15))
plot_grid(e1train2,ncol=1)

```

## Testing Phase

In Experiment 1, a single constant-trained group was compared against a
single varied-trained group. At the transfer phase, all participants
were tested from 3 positions: 1) the positions(s) from their own
training, 2) the training position(s) of the other group, and 3) a
position novel to both groups. Overall, group performance was compared
with a mixed type III ANOVA, with condition (varied vs. constant) as a
between-subject factor and throwing location as a within-subject
variable. The effect of throwing position was strong, F(3,213) = 56.12,
p\<.001, η2G = .23. The effect of training condition was significant
F(1,71)=8.19, p\<.01, η2G = .07. There was no significant interaction
between group and position, F(3,213)=1.81, p=.15, η2G = .01.

```{r}
#| label: fig-IGAS_Testing1
#| fig.cap: Testing performance for each of the 4 testing positions, compared between training conditions. Positions 610 and 910 were trained on by the varied group, and novel for the constant group. Position 760 was trained on by the constant group, and novel for the varied group. Position 835 was novel for both groups. Shorter bars are indicative of better performance (the ball landing closer to the center of the target). Error bars indicate standard error of the mean.

exp1.Test <- e1 %>% filter(stage=="Transfer") %>% select(-trainHalf)%>% group_by(positionX) %>% 
  mutate(globalAvg=mean(AbsDistFromCenter),globalSd=sd(AbsDistFromCenter)) %>% 
  group_by(sbjCode,positionX) %>% 
  mutate(scaledDev = scaleVar(globalAvg,globalSd,AbsDistFromCenter)) %>%
  ungroup() %>% group_by(sbjCode,conditType,positionX,ThrowPosition) %>%
summarise(MeanTargetDeviance = mean(AbsDistFromCenter),MeanScaleDev = mean(scaledDev),.groups="keep")%>% as.data.frame()

#manuscript plot
e1test1=exp1.Test %>% ggplot(aes(x=positionX,y=MeanTargetDeviance,group=conditType,fill=conditType))+
  geom_bar(stat="summary",fun=mean,position=dodge)+ stat_summary(fun.data=mean_se,geom="errorbar",position=dodge,width=.5)+ylab("Mean Distance From Center Of Target") +xlab("Testing Location")+theme(plot.title = element_text(hjust = 0.5))+guides(fill=guide_legend(title="Training Condition"))+theme(legend.title.align=.25)+scale_x_discrete(name="Testing Location",labels=e1Labels)

e1test1

```

\
\

```{r}
#| label: tbl-IGAS_Table1
#| tab.cap: Testing performance for varied and constant groups in experiment 1. Mean absolute deviation from the center of the target, with standard deviations in parenthesis.

exp1.Test <- e1 %>% filter(stage=="Transfer") %>% select(-trainHalf)%>% group_by(positionX) %>% 
  mutate(globalAvg=mean(AbsDistFromCenter),globalSd=sd(AbsDistFromCenter)) %>% 
  group_by(sbjCode,positionX) %>% 
  mutate(scaledDev = scaleVar(globalAvg,globalSd,AbsDistFromCenter)) %>%
  ungroup() %>% group_by(sbjCode,conditType,positionX,ThrowPosition) %>%
summarise(MeanTargetDeviance = mean(AbsDistFromCenter),MeanScaleDev = mean(scaledDev),.groups="keep")%>% as.data.frame()

test= exp1.Test %>% dplyr::rename(Condition="conditType") %>% group_by(Condition,positionX) %>%
   summarise(Mean=round(mean(MeanTargetDeviance),2),sd=round(sd(MeanTargetDeviance),2),.groups="keep")
 test=test %>% group_by(Condition) %>% mutate(GroupAvg=round(mean(Mean),2),groupSd=round(sd(Mean),2))
 test = test %>% mutate(msd=paste(Mean,"(",sd,")",sep=""),gsd=paste(GroupAvg,"(",groupSd,")",sep="")) %>% select(positionX,Condition,msd,gsd)%>%pivot_wider(names_from = Condition,values_from=c(msd,gsd))
 test=test[,1:3]

kable(test,escape=FALSE,booktabs=TRUE,col.names=c("Position","Constant","Varied"),align=c("l"))  %>% kableExtra::kable_styling(position="left") # %>%  # kable_classic() #%>% kableExtra::footnote(general=captionText,general_title = "")

```

## Discussion

In Experiment 1, we found that varied training resulted in superior
testing performance than constant training, from both a position novel
to both groups, and from the position at which the constant group was
trained, which was novel to the varied condition. The superiority of
varied training over constant training even at the constant training
position is of particular note, given that testing at this position
should have been highly similar for participants in the constant
condition. It should also be noted, though, that testing at the constant
trained position is not exactly identical to training from that
position, given that the context of testing is different in several ways
from that of training, such as the testing trials from the different
positions being intermixed, as well as a simple change in context as a
function of time. Such contextual differences will be further considered
in the General Discussion.

In addition to the variation of throwing position during training, the
participants in the varied condition of Experiment 1 also received
training practice from the closest/easiest position, as well as from the
furthest/most difficult position that would later be encountered by all
participants during testing. The varied condition also had the potential
advantage of interpolating both of the novel positions from which they
would later be tested. Experiment 2 thus sought to address these issues
by comparing a varied condition to multiple constant conditions.

# Experiment 2

In Experiment 2, we sought to replicate our findings from Experiment 1
with a new sample of participants, while also addressing the possibility
of the pattern of results in Experiment 1 being explained by some
idiosyncrasy of the particular training location of the constant group
relative to the varied group. To this end, Experiment 2 employed the
same basic procedure as Experiment 1, but was designed with six separate
constant groups each trained from one of six different locations (400,
500, 625, 675, 800, or 900), and a varied group trained from two
locations (500 and 800). Participants in all seven groups were then
tested from each of the 6 unique positions.

## Methods

### Participants

A total of 306 Indiana University psychology students participated in
Experiment 2, which was also conducted online. As was the case in
experiment 1, the undergraduate population from which we recruited
participants was 63% female and primarily composed of 18--22-year-old
individuals. Using the same procedure as experiment 1, we excluded 98
participants for exceptionally poor performance at one of the dependent
measures of the task, or for displaying a pattern of responses
indicative of a lack of engagement with the task. A total of 208
participants were included in the final analyses with 31 in the varied
group and 32, 28, 37, 25, 29, 26 participants in the constant groups
training from location 400, 500, 625, 675, 800, and 900, respectively.
All participants were compensated with course credit.

### Task and Procedure

The task of Experiment 2 was identical to that of Experiment 1, in all
but some minor adjustments to the height of the barrier, and the
relative distance between the barrier and the target. Additionally, the
intermittent testing trials featured in experiment 1 were not utilized
in experiment 2, and all training and testing trials were presented with
feedback. An abbreviated demo of the task used for Experiment 2 can be
found at (https://pcl.sitehost.iu.edu/tg/demos/igas_expt2_demo.html).

The procedure for Experiment 2 was also quite similar to experiment 1.
Participants completed 140 training trials, all of which were from the
same position for the constant groups and split evenly (70 trials each -
randomized) for the varied group. In the testing phase, participants
completed 30 trials from each of the six locations that had been used
separately across each of the constant groups during training. Each of
the constant groups thus experience one trained location and five novel
throwing locations in the testing phase, while the varied group
experiences 2 previously trained, and 4 novel locations.

## Results

### Data Processing and Statistical Packages

After confirming that condition and throwing position did not have any
significant interactions, we standardized performance within each
position, and then average across position to yield a single performance
measure per participant. This standardization did not influence our
pattern of results. As in experiment 1, we performed type III ANOVA's
due to our unbalanced design, however the pattern of results presented
below is not altered if type 1 or type III tests are used instead. The
statistical software for the primary analyses was the same as for
experiment 1. Individual learning rates in the testing phase, compared
between groups in the supplementary analyses, were fit using the TEfit
package in R [@cochraneTEfitsNonlinearRegression2020].

### Training Phase

The different training conditions trained from positions that were not
equivalently difficult and are thus not easily amenable to comparison.
As previously stated, the primary interest of the training data is
confirmation that some learning did occur. @fig-e2train depicts the training
performance of the varied group alongside that of the aggregate of the
six constant groups (5a), and each of the 6 separate constant groups
(5b). An ANOVA comparison with training stage (beginning, middle, end)
as a within-group factor and group (the varied condition vs. the 6
constant conditions collapsed together) as a between-subject factor
revealed no significant effect of group on training performance,
F(1,206)=.55,p=.49, $\eta^{2}_G$ \<.01, a significant effect of training
stage F(2,412)=77.91, p\<.001, $\eta^{2}_G$ =.05, and no significant
interaction between group and training stage, F(2,412)=.489 p=.61,
$\eta^{2}_G$ \<.01. We also tested for a difference in training
performance between the varied group and the two constant groups that
trained matching throwing positions (i.e., the constant groups training
from position 500, and position 800). The results of our ANOVA on this
limited dataset mirrors that of the full-group analysis, with no
significant effect of group F(1,86)=.48, p=.49, $\eta^{2}_G$ \<.01, a
significant effect of training stage F(2,172)=56.29, p\<.001,
$\eta^{2}_G$ =.11, and no significant interaction between group and
training stage, F(2,172)=.341 p=.71, $\eta^{2}_G$ \<.01.

```{r}

e2$stage <- factor(e2$stage, levels = c("Beginning", "Middle", "End","Transfer"),ordered = TRUE)

exp2TrainPosition <- e2  %>% filter(stage!="Transfer") %>%ungroup() %>% 
  group_by(sbjCode,Group2,conditType,trainHalf,positionX) %>% 
  summarise(MeanTargetDistance=mean(AbsDistFromCenter))%>% as.data.frame()

exp2TrainPosition3 <- e2  %>% filter(stage!="Transfer") %>%ungroup() %>% 
  mutate(globalAvg=mean(AbsDistFromCenter),globalSd=sd(AbsDistFromCenter)) %>% 
  group_by(sbjCode,positionX) %>% mutate(scaledDev = scaleVar(globalAvg,globalSd,AbsDistFromCenter)) %>%ungroup() %>%
  group_by(sbjCode,Group2,conditType,stage,positionX) %>% 
  summarise(MeanTargetDistance=mean(AbsDistFromCenter),MeanScaledDev=mean(scaledDev,trim=.05))%>% as.data.frame()

exp2Train <- e2  %>% filter(stage!="Transfer")  %>% 
  group_by(sbjCode,Group2,conditType,trainHalf) %>% 
  summarise(MeanTargetDistance=mean(AbsDistFromCenter)) %>% as.data.frame()

>>>>

---
#title: Introduction
format:
  html: default
  hugo-md:
    echo: false
    output: false
    html-math-method: mathjax
    output-file: d_intro_hugo.md
  gfm:
    echo: true
    output-file: d_into_gfm.md
---

@@
